---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# taxgraph

<!-- badges: start -->
<!-- badges: end -->

taxgraph simplifies exploration of metagenomic classifier results by parsing them into graph (tree) structures where read support are encoded as node attributes, making it easy to build sankey, sunburst, and radial phylogenetic tree visualisations. taxgraph allows visualisations to be customised by creating subgraphs based on attribute thresholds while preserving overall lineage relationships (via vertex suppression).

## Installation

You can install the development version of taxgraph from [GitHub](https://github.com/) with:

``` r
if (!require("remotes"))
    install.packages("remotes")

remotes::install_github("selkamand/taxgraph")
```

## Quick Start

taxgraph visualisations are produced in 3 steps

1) load a full taxonomy used in into a directed graph structure
2) parse metagenomic classification reports into data.frames
3) Convert tabular metagenomics report to a graph structure
3) visualise as graph


### Step 1: Parse taxonomy

These functions all return an igraph structure with the following properties

- Directed, 
- Up to 1 parent per node, multiple children per node allowed)
- Nodes with no parent allowed (we do not force a single 'root' node)
- `name` node attribute represents unique character identifier (usually taxid)
- `fullname` node attribute represents scientific name
- `rank` node attribute represents level of taxid in tree (e.g. family/genus/species). Exact terms may change depending on taxonomy type.

```{r parse_taxonomies, eval=FALSE}
library(taxgraph)

# From ktaxonomy (kraken database)
taxonomy = parse_taxonomy(system.file(package = "taxonomies/ktaxonomy.tsv"), type = "kraken")

# From taxDB file (krakenuniq database)
taxonomy = parse_taxonomy(system.file(package = "taxonomies/taxDB.tsv"), type = "taxdb")

# From nodes.dmp and names.dmp (ncbi taxonomy)
taxonomy = parse_taxonomy(system.file(package = "taxonomies/ncbi.tsv"), type = "ncbi")
```

### Step 2: Parse Metagenomic Classification Reports

These functions output a data.frame with at least the following columns

1. taxid
2. reads_covered_by_clade
3. reads_assigned_directly


```{r, eval=FALSE}

# From krakenuniq output
report_dataframe <- parse_krakenuniq_report(system.file(package = "krakenuniq.report", "reports/krakenuniq.report"))

# TODO: add kraken2 and other taxonomies
```

### Step 3: Convert tabular metagenomics report to a graph structure

Use our full taxonomy to turn the tabular report into a graph structure (where only nodes present in the report are described)

```{r, eval=FALSE}
# Create graph from all taxids in report
report_graph = report_to_graph(taxonomy, report_dataframe)
```

### Step 4: Visualise

The resulting report graph can now visualised using many different packages, but we provide some functions that make the most common plot types.

```{r, eval=FALSE}
plot_sunburst(report_graph)
plot_sankey(report_graph)
plot_radial_tree(report_graph)
```


## Similar tools

If you just want an interactive sunburst diagram we highly recommend you check out [krona](https://github.com/marbl/Krona/wiki). taxgraph was specifically designed for users who want more control over their plots, a greater range of visualisations, or prefer programmatically interacting with their taxonomies.



